---
title: "Downloading Data"
author: "Mark Padgham"
date: "`r Sys.Date()`"
#output: rmarkdown::pdf_document
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Downloading Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r load, eval=FALSE}
#library (osmplotr)
devtools::install_github ('mpadge/osmplotr')
```
```{r, echo=FALSE, message=FALSE}
require (devtools)
setwd ("../..")
load_all ("osmplotr")
setwd ("./osmplotr/vignettes")
```

This vignette demonstrates how the `london` data included with the `osmplotr`
package were generated. First customise an `osm_structures` data frame.

```{r}
# better with a bigger bbox:
# bbox <- get_bbox (c(-0.15,51.5,-0.1,51.52))
bbox <- get_bbox (c(-0.13,51.5,-0.11,51.52))
structures <- c ('highway', 'highway', 'building', 'building', 'building',
                 'amenity', 'grass', 'park', 'natural', 'tree')   
structs <- osm_structures (structures=structures, col_scheme='dark')   
structs$value [1] <- '!primary'   
structs$value [2] <- 'primary'
structs$suffix [2] <- 'HP'
structs$value [3] <- '!residential'
structs$value [4] <- 'residential'
structs$value [5] <- 'commercial'
structs$suffix [3] <- 'BNR'
structs$suffix [4] <- 'BR'
structs$suffix [5] <- 'BC'
```

Then download the corresponding data, noting that `extract_osm_objects` returns
two components: `warn` containing any warnings generated during download, and
`obj` containing the appropriate spatial object.

```{r, eval=FALSE}
london <- list ()
for (i in 1:(nrow (structs) - 1)) 
{
    dat <- extract_osm_objects (key=structs$key [i], value=structs$value [i],
                                bbox=bbox)
    if (!is.null (dat$warn))
        warning (dat$warn)
    fname <- paste0 ('dat_', structs$suffix [i])
    assign (fname, dat$obj)
    london [[i]] <- get (fname)
    names (london)[i] <- fname
    rm (list=c(fname))
}
```

And finally the additional data for specific buildings and highways (ignoring
potential warnings this time).

```{r, eval=FALSE}
extra_pairs <- c ('name', 'Royal.Festival.Hall')
london$dat_RFH <- extract_osm_objects (key='building', extra_pairs=extra_pairs, 
                                       bbox=bbox)$obj
extra_pairs <- list (c ('addr:street', 'Stamford.St'),
                     c ('addr:housenumber', '150'))
london$dat_ST <- extract_osm_objects (key='building', extra_pairs=extra_pairs, 
                                      bbox=bbox)$obj
# These examples need expanded bboxes
bbox <- get_bbox (c(-0.15,51.5,-0.10,51.52))
highways <- c ('Kingsway', 'Holborn', 'Farringdon.St', 'Strand',
               'Fleet.St', 'Aldwych')
london$highways1 <- highways2polygon (highways=highways, bbox=bbox)
highways <- c ('Queen.s.Walk', 'Blackfriars', 'Waterloo', 'The.Cut')
london$highways2 <- highways2polygon (highways=highways, bbox=bbox)
highways <- c ('Regent.St', 'Oxford.St', 'Shaftesbury')
london$highways3 <- highways2polygon (highways=highways, bbox=bbox)
```
